---
title: "STUDY122 / BAN2401-301 - Data processing script"
author: "C2N Diagnostics - proprietary and confidential information. Not to be copied or distributed without C2N written consent."
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)

now <<- Sys.time()

library(rmarkdown)
library(ggplot2)
library(pROC)
library(data.table)
library(caret)
library(plyr)
library(dplyr)
library(ggpubr)
library(expss)
library(ggforce)
library(gridExtra)
# visualizing the effects of variables in a GLM
library(effects)
library(Matrix)

# For nicer looking html tables
library(formattable)
library(readxl)

# For summary tables
library(qwraps2)
options(qwraps2_markup = "markdown")

library(DiagrammeR)
#http://rich-iannone.github.io/DiagrammeR/graphviz_and_mermaid.html
#https://bookdown.org/yihui/rmarkdown-cookbook/diagrams.html




# Plot & visualization
theme_new <- function(base_size = 12, base_family = ""){
  theme_bw(base_size = base_size, base_family = base_family) %+replace%
    theme(
      legend.key=element_rect(colour=NA, fill =NA),
      panel.grid = element_blank(),   
      panel.border = element_rect(fill = NA, colour = "black", size=1),
      panel.background = element_rect(fill = "white", colour = "black"), 
      strip.background = element_rect(fill = NA)
    )
}

theme_set(theme_new(base_size = 16))


printpct = function(percentage) {
  percentage = as.numeric(percentage)
  
  if (is.na(percentage) | percentage == 0)
    return("-")
  else
    return(paste0(percentage, "%"))
}

categorize_aps = function (APS) {
  
  if (is.na(APS)) {
    return(NA)
  }
  
  if (APS <= 35) {
    return("Low")
  } else if (APS >= 58) {
    return("High")
  } else {
    return("Intermediate")
  }
  
}


## date given as number of days since 1900-01-01 (a date in 1989)
##  as.Date(32768, origin = "1900-01-01")
## Excel is said to use 1900-01-01 as day 1 (Windows default) or
## 1904-01-01 as day 0 (Mac default), but this is complicated by Excel
## incorrectly treating 1900 as a leap year.
## So for dates (post-1901) from Windows Excel:
##  as.Date(35981, origin = "1899-12-30") # 1998-07-05
getDateFromExcel = function (datestring) {
  
  converteddate = NA
  
  if(is.na(as.numeric(datestring))) {
    # the date is a string not a number
    converteddate = as.Date(datestring, tryFormats = c("%d%b%Y"))
    
  } else {
    # the date is stoed as Excel date - which is converted to a number when casta s text
    converteddate = as.Date(as.numeric(datestring), origin = "1899-12-30")
    
  }
  
  return(as.character(converteddate))
  
}

```


# Put data together

## Load in the sample list from Hongmei


Load in Batch 1&2

```{r}
#hongmeidata = read_xlsx("Batch1 Shipment_C2N_updated_13Jul2022.xlsx", col_types = c("text"))  
hongmeidata12 = read_xlsx("Batch2_Shipment.xlsx", col_types = c("text"))  

```

Load in Batch 3&4

```{r}

hongmeidata34 = read_xlsx("C2N Batch 3&4 Sample list _18Aug2022_sent (original from Hongmei.xlsx", col_types = c("text"))

```


Colnames differences

```{r}

setdiff(colnames(hongmeidata12), colnames(hongmeidata34))
setdiff(colnames(hongmeidata34), colnames(hongmeidata12))

```

Keep just what we need from both

```{r}

hongmeidata12 = hongmeidata12[, c("Specimen_ID", "Sampled_Date", "Screen_Id", "Protocol", "Visit_Name", "Visit_Description")]
hongmeidata34 = hongmeidata34[, c("Specimen_ID", "Sampled_Date", "Screen_Id", "Protocol", "Visit_Name", "Visit_Description")]

hongmeidata = hongmeidata34 #rbind(hongmeidata12, hongmeidata34)

#ids = rbind(hongmeidata[duplicated(hongmeidata$Specimen_ID), ]
#ids = ids[order(ids$Specimen_ID), ]

```





```{r}

# Fix the date
hongmeidata$PCDTC = as.character(as.Date(substr(hongmeidata$Sampled_Date, 1, 7), format = "%d%b%y"))

# Fix study ID
hongmeidata[which(hongmeidata$Protocol == "BAN2401-G000-301"), "STUDYID"] = "BAN2401 G000 301"

hongmeidata_c = hongmeidata[, c("Specimen_ID", "PCDTC", "Screen_Id", "STUDYID", "Visit_Name", "Visit_Description")]

colnames(hongmeidata_c) = c("PCREFID", "PCDTC", "SUBJID", "STUDYID", "VISITNAME", "VISITDESCRIPTION")

table(hongmeidata_c$VISITDESCRIPTION)

```

### Translate VISITNAME to VISIT

```{r}

#    VISITNAME == "V22" ~ "UNSCH",
#    VISITNAME == "V6" ~ "UNSCH",
#    VISITNAME == "V9" ~ "UNSCH",


hongmeidata_c = hongmeidata_c %>% 
  mutate(VISIT = case_when(
# This is a case statement - so do the VISITDESCRIPTION unscheduled ones first
    VISITDESCRIPTION == "UNSCH / RETEST" ~ "UNSCH",
    VISITDESCRIPTION == "AMD6 UNSCH/RETEST" ~ "UNSCH",
    
# Then look at the VISITNAME
    VISITNAME == "V3W1" ~ "V3W1",
    VISITNAME == "V3" ~ "V3W1",
    VISITNAME == "V16" ~ "V16W27",
    VISITNAME == "V29" ~ "V29W53",
    VISITNAME == "V41" ~ "V41W77",
    VISITNAME == "V41 AMD6" ~ "V41W77",
    VISITNAME == "Early Term" ~ "ET",
    VISITNAME == "EARLY TERM" ~ "ET",
    VISITNAME == "EARLY TERM AMEND 01" ~ "ET",
    VISITNAME == "Early Term AMEND01" ~ "ET",
    VISITNAME == "V42" ~ "V42W79",
    VISITNAME == "V42A" ~ "V42W79",
    VISITNAME == "V50" ~ "V50W93",
    VISITNAME == "V56" ~ "V56W105",
    VISITNAME == "V42 AMD6" ~ "V42W79",
    VISITNAME == "V50 AMD6" ~ "V50W93",
    VISITNAME == "V56 AMD6" ~ "V56W105",
    VISITNAME == "V70" ~ "V70W133",
    VISITNAME == "V82" ~ "V82W157",
    VISITNAME == "V93" ~ "V93W179",

# Then default value
    TRUE ~ ""
  )
)

#table(hongmeidata_c$VISIT)
#table(manifest_c$VISITNAME)

# touble subject: 10541038

hongmeidata_c[which(hongmeidata_c$SUBJID == "10541038"), ]


```

How did that go?

```{r}

paste0("We have ", length(hongmeidata_c$PCREFID), " entries in the full list")
paste0("Of those ", length(hongmeidata_c[which(hongmeidata_c$VISIT == ""), ]$PCREFID), " entries did not match the right visit type")

table(hongmeidata_c[which(hongmeidata_c$VISIT == ""), ]$VISITNAME)

```


First 6 entries:

```{r}

head(hongmeidata_c)

```


### Information from manifest

The following information comes from the external file:


| Field Name | Value       |
|------------|-------------|
| SUBJID    | from manifest     |
| VISIT      | from manifest     |
| PCDTC      | from manifest   |
| STUDYID    | from manifest    |



### Unique SUBJID

```{r}

paste0("Out of **", length(hongmeidata_c$SUBJID), "** entries in the manifest there are ", length(unique(hongmeidata_c$SUBJID)), " unique subject identifiers.")

```




## Load in the cumulative datafile

```{r}

# JOSH CHANGE THE DATE TO THE DATE IN THE FILENAME LIKE
# All Abeta Data_23MAR2022.xlsx
#dateoffile = "24Aug2022"
#dateoffile = "14Sep2022"

# Email from Venky, 04 OCT 2022. There are additional 11 samples that have data. Data was sent from Kris. I (TW) had to manually add this data to the end of the data file from 14Sept2022 to include this data in a new file:
dateoffile = "04Oct2022"

# Read Abeta file and force column types
abetadata = read_xlsx(paste0("cumulativefiles/STUDY122_Abeta_Data_", dateoffile, ".xlsx"), 
                      col_types = c("text", "numeric", "numeric", "text", "text", "text", "numeric", "text", "numeric", "date"))

# Rename columns per DTA - for this they use PC

colnames(abetadata) = c("PCREFID", "Abeta40_Plasma_IPMS", "Abeta42_Plasma_IPMS", "Abeta4240_Plasma_IPMS_excel", "Abeta_comment", "PCDTC_excel", "YearOfBirth", "SUBJID_excel", "batch", "rundate")

abetadata$Abeta4240_Plasma_IPMS = abetadata$Abeta42_Plasma_IPMS / abetadata$Abeta40_Plasma_IPMS

```

### Data points per batch

Abeta

```{r}

table(abetadata$batch)

```


### Number of unique sample identifiers

Abeta file unique IDs:

```{r}

length(unique(abetadata$PCREFID))

```


```{r}

if (length(unique(abetadata$PCREFID)) != length(abetadata$PCREFID)) {
  
  print(paste0("The number of unique PCREFIDs (", length(unique(abetadata$PCREFID)), ") is not the same as the number of data points (", length(abetadata$PCREFID), ")"))
  print("Before we merge the data we must take out any duplicate data from the Abeta data files.")
  print("The following batches contain duplicated PCREFIDs: ")

  print(unique(abetadata[duplicated(abetadata$PCREFID), c("PCREFID", "batch")]))
  print(unique(abetadata[duplicated(abetadata$PCREFID, fromLast = T), c("PCREFID", "batch")]))
  
  knitr::knit_exit()
  
} else {
  
  print("No duplicate PCREFIDs in the Abeta data file")
  
}


```


## Merge data from cumulative file with region files

```{r}

study122data = merge(abetadata, hongmeidata_c, by = c("PCREFID"), all.x = T)

```

### Check on merge


```{r}

paste0("Out of a total of ", length(study122data$PCREFID), " data points, ", sum(as.numeric(!is.na(study122data$SUBJID))), " samples got a subject ID from the manifest")



if(sum(as.numeric(is.na(study122data$SUBJID))) > 0) {

  print(paste0("There are ", sum(as.numeric(is.na(study122data$SUBJID))), " patients in the datafile that are not present in the manifest - these should be reconciled."))  
  formattable(study122data[which(is.na(study122data$SUBJID)), c("PCREFID", "SUBJID", "VISIT", "PCDTC", "batch", "rundate")])

  print(table(study122data[which(is.na(study122data$SUBJID)), "batch"]))
  
  
#  knitr::knit_exit()
  
}


```



### Make sure all have valid VISIT information

```{r}

#manifest_i = manifest_i[complete.cases(manifest_i[ , c("Protocol", "Patient ID")]),]

if (length(which(study122data$VISIT == "")) > 0) {

  print(paste0(length(which(study122data$VISIT == "")), " VISITNAMEs were not translated to VISIT - either a wrong sample was analyzed or the VISITNAME was typed wrong"))
  
  print(formattable(study122data[which(study122data$VISIT == ""), c("PCREFID", "SUBJID", "VISIT", "VISITNAME")]))
  
  print(table(study122data[which(study122data$VISIT == ""), c("VISITNAME")]))
  
#  knitr::knit_exit()
  
} else {
  
  print("All VISITNAMEs translated")
}



```



### SubjectIDs 

Comes from the manifest. Empty SUBJID:

```{r}

table(is.na(study122data$SUBJID))


```

Unique SUBJID:

```{r}

length(unique(study122data$SUBJID))

```



### Are subjects analyzed on one or more batches?

```{r}

#table(study122data$SUBJID, study122data$batch)

subjectids = unique(study122data$SUBJID)

subjectid = "10011007"

multiplebatches = data.frame()

for (subjectid in subjectids) {
  
  dataset = study122data[which(study122data$SUBJID == subjectid), ]
  batches = unique(dataset$batch)
  
  if (length(batches) > 1) {
    #print(paste0("Samples from subject ", subjectid, " is present on more than one batch (", paste(batches, collapse= ", "), ")"))

  }
  
  multiplebatches = rbind(multiplebatches, data.frame("SUBJID" = subjectid, "numberofbatches" = length(batches)))
  
}

paste0("Number of patients with X number of batches")

table(multiplebatches$numberofbatches)

```





## Clean out any samples with comments

```{r}

study122data$Abeta_comment[which(study122data$Abeta_comment == "")] = NA

```

### Abeta comments:

```{r}

formattable(table(study122data$Abeta_comment))

```


Change "\u03b2" to "beta"

```{r}

study122data$Abeta_comment = gsub('\u03b2', 'beta', study122data$Abeta_comment)

formattable(table(study122data$Abeta_comment))

```



### Samples with comments = Not Done

For all samples with sample comments:

* Set the results to NA
* Set status to ND

Set the reason not done to a concatenation of Abeta comments and then make sure that the reason not done it not longer than 200 characters as per the DTA.

```{r}

study122data$PCSTAT = ""

sampleswithqns = which(!is.na(study122data$Abeta_comment))

paste0("Number of samples: ", length(study122data$PCREFID), ", samples with comments: ", length(sampleswithqns))

if (length(sampleswithqns) > 0) {
  study122data[sampleswithqns, c("Abeta40_Plasma_IPMS", "Abeta42_Plasma_IPMS", "Abeta4240_Plasma_IPMS")] = NA
  
  # Set status and reason not done
  study122data[sampleswithqns, c("PCSTAT")] = "ND"
  
}



# convert NA back to ""
study122data$Abeta_comment[is.na(study122data$Abeta_comment)] = ""

# use the comment as the reason not done
study122data$PCREASND = study122data$Abeta_comment

# make sure that reason not done does not overflow the 200 characters
study122data$PCREASND = substr(study122data$PCREASND,1,200)

table(study122data$PCREASND)

#study122data[sampleswithqns, c("PCREASND")] = "QNS"

```



# Data review

```{r}

# EXPORT DATA
studyid = "STUDY122"
modeldata = study122data[, c("Abeta42_Plasma_IPMS", "Abeta40_Plasma_IPMS", "Abeta4240_Plasma_IPMS")]

modeldata$race = NA
modeldata$ethnicity = NA
modeldata$gender = NA
modeldata$id = paste0(studyid, "_", study122data$PCREFID)
modeldata$cohort = studyid
#visual read positive - in this case not VR 
modeldata$amyloidpositive = NA
modeldata$centiloid = NA
modeldata$MMSE = NA
modeldata$CDR = NA
modeldata$tracer = NA
modeldata$diagnosis_desc = NA
modeldata$state = "NA"

write.csv(modeldata, paste0("small_model_data_", studyid, ".csv"))


```



## Demographics

```{r results='asis'}


demographics_table = 
  list(

        "Blood Aβ42/40" = 
    list("mean (sd)" = ~ qwraps2::mean_sd(na.omit(.data$Abeta4240_Plasma_IPMS), denote_sd = "paren", digits = 3),
         "min" = ~ round(min(na.omit(.data$Abeta4240_Plasma_IPMS)), 3),
         "max" = ~ round(max(na.omit(.data$Abeta4240_Plasma_IPMS)), 3)
         ),
    
        "Blood Aβ40" = 
    list("mean (sd)" = ~ qwraps2::mean_sd(na.omit(.data$Abeta40_Plasma_IPMS), denote_sd = "paren", digits = 1),
         "min" = ~ round(min(na.omit(.data$Abeta40_Plasma_IPMS)), digits = 1),
         "max" = ~ round(max(na.omit(.data$Abeta40_Plasma_IPMS)), digits = 1)
         ),
    
        "Blood Aβ42" = 
    list("mean (sd)" = ~ qwraps2::mean_sd(na.omit(.data$Abeta42_Plasma_IPMS), denote_sd = "paren", digits = 1),
         "min" = ~ round(min(na.omit(.data$Abeta42_Plasma_IPMS)), digits = 1),
         "max" = ~ round(max(na.omit(.data$Abeta42_Plasma_IPMS)), digits = 1)
         )
        
  )




# Show data for all subjects
c = summary_table(study122data, demographics_table)

colnames(c) = gsub("x", "All", colnames(c))

c

```



## Abeta histograms

```{r}
ggplot(study122data, aes(x=Abeta4240_Plasma_IPMS)) + 
  geom_histogram(aes(y=..density..),fill="dodgerblue3",color="white",alpha=0.7, binwidth = 0.005) + 
  geom_density() +
  geom_rug() +
  geom_vline(xintercept = 0.089) +
  labs(x='Abeta4240_Plasma_IPMS') +
  annotate("text", x=mean(study122data$Abeta4240_Plasma_IPMS), y=20, label = paste0("mean = ", round(mean(study122data$Abeta4240_Plasma_IPMS), 3))) +
  theme_minimal()

ggplot(study122data, aes(x=Abeta40_Plasma_IPMS)) + 
  geom_histogram(aes(y=..density..),fill="dodgerblue3",color="white",alpha=0.7, binwidth = 50) + 
  geom_density() +
  geom_rug() +
  labs(x='Abeta40_Plasma_IPMS') +
  annotate("text", x=mean(study122data$Abeta40_Plasma_IPMS), y=0.0025, label = paste0("mean = ", round(mean(study122data$Abeta40_Plasma_IPMS), 3))) +
  theme_minimal()

ggplot(study122data, aes(x=Abeta42_Plasma_IPMS)) + 
  geom_histogram(aes(y=..density..),fill="dodgerblue3",color="white",alpha=0.7, binwidth = 5) + 
  geom_density() +
  geom_rug() +
  labs(x='Abeta42_Plasma_IPMS') +
  annotate("text", x=mean(study122data$Abeta42_Plasma_IPMS), y=0.025, label = paste0("mean = ", round(mean(study122data$Abeta42_Plasma_IPMS), 3))) +
  theme_minimal()


```


## Abeta averages tracking by batch

```{r}

ggplot(study122data, aes(x=batch, y=Abeta4240_Plasma_IPMS)) + 
  geom_point() +
  geom_smooth(method = loess) +
  labs(y='Abeta4240_Plasma_IPMS', x = "Batch") +
  theme_minimal()

ggplot(study122data, aes(x=batch, y=Abeta40_Plasma_IPMS)) + 
  geom_point() +
  geom_smooth(method = loess) +
  labs(y='Abeta40_Plasma_IPMS', x = "Batch") +
  theme_minimal()

ggplot(study122data, aes(x=batch, y=Abeta42_Plasma_IPMS)) + 
  geom_point() +
  geom_smooth(method = loess) +
  labs(y='Abeta42_Plasma_IPMS', x = "Batch") +
  theme_minimal()

```


## Abeta averages tracking by date

```{r}

ggplot(study122data, aes(x=rundate, y=Abeta4240_Plasma_IPMS, color = as.factor(batch))) + 
  geom_boxplot() +
  labs(y='Abeta4240_Plasma_IPMS', x = "Date") +
  theme_minimal()

ggplot(study122data, aes(x=rundate, y=Abeta4240_Plasma_IPMS)) + 
  geom_point() +
  geom_smooth(method = loess) +
  labs(y='Abeta4240_Plasma_IPMS', x = "Date") +
  theme_minimal()

ggplot(study122data, aes(x=rundate, y=Abeta40_Plasma_IPMS)) + 
  geom_point() +
  geom_smooth(method = loess) +
  labs(y='Abeta40_Plasma_IPMS', x = "Date") +
  theme_minimal()

ggplot(study122data, aes(x=rundate, y=Abeta42_Plasma_IPMS)) + 
  geom_point() +
  geom_smooth(method = loess) +
  labs(y='Abeta42_Plasma_IPMS', x = "Date") +
  theme_minimal()

```

## Longitudinal data


### Abeta 40

```{r fig.height=18, fig.width=14}

ggplot(study122data, aes(x=as.Date(PCDTC), y=Abeta40_Plasma_IPMS, color = as.factor(SUBJID)))+
  geom_point() +
  facet_wrap(~ batch) +
  geom_line(size = 0.5) +
  theme(legend.position = "none") +
  labs(title= NULL,
           x = "Date of sample collection", 
           y = "Abeta40 (pg/mL)")

```


### Abeta 42


```{r fig.height=18, fig.width=14}

ggplot(study122data, aes(x=as.Date(PCDTC), y=Abeta42_Plasma_IPMS, color = as.factor(SUBJID)))+
  geom_point() +
  facet_wrap(~ batch) +
  geom_line(size = 0.5) +
  theme(legend.position = "none") +
  labs(title= NULL,
           x = "Date of sample collection", 
           y = "Abeta42 (pg/mL)")

```

### Abeta Ratio


```{r fig.height=18, fig.width=14}

ggplot(study122data, aes(x=as.Date(PCDTC), y=Abeta4240_Plasma_IPMS, color = as.factor(SUBJID)))+
  geom_point() +
  geom_hline(yintercept = 0.089) +
  facet_wrap(~ batch) +
  geom_line(size = 0.5) +
  theme(legend.position = "none") +
  labs(title= NULL,
           x = "Date of sample collection", 
           y = "Abeta42/40 (no unit)")

```


### Longitudinal Ratio by Visit

Visits in the data set:

```{r}

table(study122data$VISIT)

```

```{r fig.height=18, fig.width=14}


level_order = c("V3W1", "V16W27", "V29W53", "V41W77", "V42W79", "ET", "")

ggplot(study122data, aes(x=factor(VISIT, level = level_order), y=Abeta4240_Plasma_IPMS, color = as.factor(SUBJID)))+
  geom_point() +
  geom_hline(yintercept = 0.089) +
  facet_wrap(~ batch) +
  geom_line(size = 0.5) +
  theme(legend.position = "none") +
  labs(title= NULL,
           x = "Date of sample collection", 
           y = "Abeta42/40 (no unit)")


```





# Get data ready for export

```{r}
currentDate = format(Sys.Date(), "%Y%m%d") 


# For test transfer - x has to be replaced with a transfer increment
# ** TEST TRANSFER SHOULD ONLY BE DONE WITH DUMMY DATA TO AVOID UNBLINDING **
increment_txt = "testx"

# For production transfer - x has to be replaced with a transfer increment
increment_txt = "x"

# For database lock transfer 
increment_txt = "dblock"



filename = paste0("BAN2401_G000_301_AB4042_C2N_", currentDate, "_", increment_txt, ".CSV")
filename_for_file = paste0("BAN2401_G000_301_AB4042_C2N_", currentDate, "_", ".CSV")


```


Filename:

**`r filename `**

## Convert data from wide to long format


```{r}

# 1) Keep just the data we report
study122data.export = study122data[, c("STUDYID", "SUBJID", "PCREFID", "PCSTAT", "PCREASND", "PCDTC", "VISIT", "Abeta40_Plasma_IPMS", "Abeta42_Plasma_IPMS", "Abeta4240_Plasma_IPMS")]
colnames(study122data.export) = c("STUDYID", "SUBJID", "PCREFID", "PCSTAT", "PCREASND", "PCDTC", "VISIT", "AB40", "AB42", "AB4240")

# 1.1) Round the ratio to 3 sig digits
study122data.export$AB4240 = round(study122data.export$AB4240, 9)

# 1.2) convert all results to text - we want to report lack of result as "" - this only works if we have the result as text string
study122data.export$AB40 = as.character(study122data.export$AB40)
study122data.export$AB42 = as.character(study122data.export$AB42)
study122data.export$AB4240 = as.character(study122data.export$AB4240)


# 2) Convert to long format
study122data.export.long = melt(study122data.export, id.vars = c("STUDYID", "SUBJID", "PCREFID", "PCSTAT", "PCREASND", "PCDTC", "VISIT"), variable.name = "PCTESTCD", value.name = "PCORRES")

# 3) Sort by SUBJID & PCREFID
study122data.export.long = study122data.export.long[
  with(study122data.export.long, order(SUBJID, PCREFID)),
]




# For QNS samples all the results are NA - but they do not want NA in the PCORRES column so we will change all NA to ""
#study122data.export.long[which(is.na(study122data.export.long$PCORRES)), "PCORRES"] = "" 
study122data.export.long[which(is.na(study122data.export.long$PCORRES)), ]$PCORRES = "" 
  

formattable(sample_n(study122data.export.long, 20))


```

## Add the other fields to the export file

Order of fields in the output file:

| Field Name | Value       |
|------------|-------------|
| STUDYID    | from manifest    |
| DOMAIN     | "PC"        |
| SITE       | first 4 characters of PCREFID |
| SUBJID    | from manifest     |
| PCNAM      | "C2N"       |
| PCREFID    | dynamic     |
| VISIT      | from manifest     |
| PCDTC      | from manifest   |
| PCSPEC     | "PLASMA"    |
| PCCAT      | "BIOMARKER" |
| PCTESTCD   | dynamic     |
| PCORRES    | dynamic     |
| PCORRESU   | per above   |
| PCMETHOD   | "LC-MS/MS"  |
| PCSTAT     | dynamic     |
| PCREASND   | dynamic     |
| PCTEST     | dynamic based on above, see table below |
| PCLLOQ     | based on analyte |
| PCULOQ     | based on analyte |
| PCSPCCND   | ""          | 
| FILENAME   | static = filename_for_file |



Test codes and test names:

+---------------+------------------------------------+----------------+
| PCTESTCD      | PCTEST                             | PCORRESU       |
+===============+====================================+================+
| AB42          | C2N Abeta 42 concentration         | pg/mL          |
+---------------+------------------------------------+----------------+
| AB40          | C2N Abeta 40 concentration         | pg/mL          |
+---------------+------------------------------------+----------------+
| AB4240        | C2N Abeta 42/40 ratio              |                |
+---------------+------------------------------------+----------------+



```{r}
# STATIC ONES FIRST:
study122data.export.long$DOMAIN   = "PC"
#study122data.export.long$SITE     = ""
study122data.export.long$PCNAM    = "C2N" 
study122data.export.long$PCSPEC   = "PLASMA"  
study122data.export.long$PCCAT    = "BIOMARKER"
study122data.export.long$PCMETHOD = "LC-MS/MS"
study122data.export.long$PCSPCCND = ""
study122data.export.long$FILENAME = filename_for_file


# SITEID is the first 4 characters of SUBJID
study122data.export.long$SITE = substr(study122data.export.long$SUBJID , 1, 4)

# UNIT DEPENDS ON PCTESTCD
study122data.export.long$PCORRESU = "" 
study122data.export.long$PCORRESU[which(study122data.export.long$PCTESTCD == "AB42" | study122data.export.long$PCTESTCD == "AB40")] = "pg/mL"

# PCTEST depends on PCTESTCD
study122data.export.long = study122data.export.long %>% 
  mutate(PCTEST = case_when(
    .$PCTESTCD == "AB42" ~ "C2N Abeta 42 concentration",
    .$PCTESTCD == "AB40" ~ "C2N Abeta 40 concentration",
    .$PCTESTCD == "AB4240" ~ "C2N Abeta 42/40 ratio",
    TRUE ~ ""
  )
)


# PCLLOQ depends on PCTESTCD
study122data.export.long = study122data.export.long %>% 
  mutate(PCLLOQ = case_when(
    .$PCTESTCD == "AB42" ~ "2",
    .$PCTESTCD == "AB40" ~ "11",
    TRUE ~ ""
  )
)

# PCULOQ depends on PCTESTCD
study122data.export.long = study122data.export.long %>% 
  mutate(PCULOQ = case_when(
    .$PCTESTCD == "AB42" ~ "3920",
    .$PCTESTCD == "AB40" ~ "26400",
    TRUE ~ ""
  )
)


# WE ALSO HAVE TO IMPLEMENT THESE LIMITS:
# 1) convert all results to character but keep as values in study122data.export.long$PCORRES_NUM
study122data.export.long$PCORRES_NUM = as.numeric(study122data.export.long$PCORRES)
#study122data.export.long$PCORRES = as.character(study122data.export.long$PCORRES)

study122data.export.long[which((study122data.export.long$PCTESTCD == "AB42" & study122data.export.long$PCORRES_NUM < 2) | 
                               (study122data.export.long$PCTESTCD == "AB40" & study122data.export.long$PCORRES_NUM < 11)), "PCORRES"] = "BLQ"

study122data.export.long[which((study122data.export.long$PCTESTCD == "AB42" & study122data.export.long$PCORRES_NUM > 3920) | 
                               (study122data.export.long$PCTESTCD == "AB40" & study122data.export.long$PCORRES_NUM > 26400)), "PCORRES"] = "ALQ"


sampleswithinvalidratios = unlist(study122data.export.long[which((study122data.export.long$PCORRES == "BLQ") | (study122data.export.long$PCORRES == "ALQ")), "PCREFID"])

study122data.export.long[which(study122data.export.long$PCREFID %in% sampleswithinvalidratios & study122data.export.long$PCTESTCD == "AB4240"), "PCSTAT"] = "ND"
study122data.export.long[which(study122data.export.long$PCREFID %in% sampleswithinvalidratios & study122data.export.long$PCTESTCD == "AB4240"), "PCREASND"] = "Ratio cannot be calculated"
study122data.export.long[which(study122data.export.long$PCREFID %in% sampleswithinvalidratios & study122data.export.long$PCTESTCD == "AB4240"), "PCORRES"] = NA
study122data.export.long[which(study122data.export.long$PCREFID %in% sampleswithinvalidratios & study122data.export.long$PCTESTCD == "AB4240"),]


```

Sites:

```{r}
table(study122data.export.long$SITE)
```


```{r}
# ORDER THE COLUMNS


col_order = c("STUDYID", 
              "SITE", 
              "SUBJID", 
              "VISIT",
              "DOMAIN", 
              "PCCAT",
              "PCDTC",
              "PCREFID", 
              "PCNAM", 
              "PCTESTCD",
              "PCTEST",
              "PCORRES",
              "PCORRESU",
              "PCLLOQ",
              "PCULOQ",
              "PCMETHOD",
              "PCSPEC",
              "PCSPCCND",
              "PCSTAT",
              "PCREASND",
              "FILENAME"
              )

# Should be 21



```


## Depending on type of transfer - clear the results

If not dblock or testx (is: `r increment_txt `) - then clear:

* PCORRES
* PCSTAT
* PCREASND

Per DTA we will only provide these variables in test transfer and in dblock transfer.

```{r}

nullvalue = ""

if (increment_txt != "dblock" & increment_txt != "testx") {
  
  study122data.export.long$PCORRES = nullvalue
  study122data.export.long$PCSTAT = nullvalue
  study122data.export.long$PCREASND = nullvalue
  
}

paged_table(head(study122data.export.long))

```

## Sorted data for export


```{r}
study122data.export.long.sort = subset(study122data.export.long, select = col_order)

paged_table(study122data.export.long.sort)
  
```



Number of samples in final file:

```{r}

length(study122data$PCREFID)

```

Number of lines in the long export file divided by 3 since there are 3 results per sample.

```{r}

length(study122data.export.long.sort$SUBJID)/3

```



```{r message=FALSE}

## TEST
# BAN2401_TEST_C2N_BM_FULL_YYYYMMDDX.csv f

## PROD
# BAN2401_PROD_C2N_BM_FULL_YYYYMMDDX.csv 

write.csv(study122data.export.long.sort, filename, row.names=FALSE)

message(paste0("Data export saved to: ", filename))
message(paste0("Time to run this code: ", difftime(Sys.time(), now)))

```





