---
title: "STUDY138 - Interim Unaudited Analysis"
author: "C2N Diagnostics - proprietary and confidential information. Not to be copied or distributed without C2N written consent."
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true

---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)

now <<- Sys.time()

library(rmarkdown)
library(ggplot2)
library(pROC)
library(data.table)
library(caret)
library(plyr)
library(dplyr)
library(ggpubr)
library(expss)
library(ggforce)
library(gridExtra)
# visualizing the effects of variables in a GLM
library(effects)
library(Matrix)
library(readxl)

# For nicer looking html tables
library(formattable)

# For summary tables
library(qwraps2)
options(qwraps2_markup = "markdown")

library(DiagrammeR)
#http://rich-iannone.github.io/DiagrammeR/graphviz_and_mermaid.html
#https://bookdown.org/yihui/rmarkdown-cookbook/diagrams.html




# Plot & visualization
theme_new <- function(base_size = 12, base_family = ""){
  theme_bw(base_size = base_size, base_family = base_family) %+replace%
    theme(
      legend.key=element_rect(colour=NA, fill =NA),
      panel.grid = element_blank(),   
      panel.border = element_rect(fill = NA, colour = "black", size=1),
      panel.background = element_rect(fill = "white", colour = "black"), 
      strip.background = element_rect(fill = NA)
    )
}

theme_set(theme_new(base_size = 16))


printpct = function(percentage) {
  percentage = as.numeric(percentage)
  
  if (is.na(percentage) | percentage == 0)
    return("-")
  else
    return(paste0(percentage, "%"))
}

categorize_aps = function (APS) {
  
  if (is.na(APS)) {
    return(NA)
  }
  
  if (APS <= 35) {
    return("Low")
  } else if (APS >= 58) {
    return("High")
  } else {
    return("Intermediate")
  }
  
}

```

# Load the data

```{r}



```



## Mass spec data

```{r }

# JOSH CHANGE THIS NUMBER:
maxfile = 34

# HOW MANY FILES TO IMPORT
filenos = seq(2, maxfile, 1)
#filenos = c(filenos, 12)

abetadata = read_excel("biobankedsamples/Study138_Abeta_Results_Run_1.xlsx")
abetadata$batch = 1

apoedata  = read_excel("biobankedsamples/Study138_ApoE_Results_Run_1.xlsx")

samplesperfile = data.frame("Batch" = 1, "Abeta" = length(abetadata$`Plasma Barcodes`), "ApoE" = length(apoedata$`Plasma Barcodes`))

for (fileno in filenos) {

  newabetadata = read_excel(paste0("biobankedsamples/Study138_Abeta_Results_Run_", fileno, ".xlsx"))
  newabetadata$batch = fileno
  abetadata = rbind(abetadata, newabetadata)
  
  newapoedata = read_excel(paste0("biobankedsamples/Study138_ApoE_Results_Run_", fileno, ".xlsx"))
  apoedata  = rbind(apoedata, newapoedata)
  
  samplesperfile = rbind(samplesperfile, data.frame("Batch" = fileno, "Abeta" = length(newabetadata$`Plasma Barcodes`), "ApoE" = length(newapoedata$`Plasma Barcodes`)))

}

colnames(abetadata) = c("LBREFID", "Abeta40_Plasma_IPMS", "Abeta42_Plasma_IPMS", "Abeta4240_Plasma_IPMS_excel", "Abeta_comment", "batch")
colnames(apoedata) = c("LBREFID", "Proteotype", "ApoE_comment")

#length(unique(abetadata$LBREFID))
#length(unique(apoedata$LBREFID))

study138data.c2n = merge(abetadata, apoedata)

```

Number of samples per batch for Abeta and ApoE

```{r}
formattable(samplesperfile)
```

Total number of samples

```{r}

length(study138data.c2n$LBREFID)

```

## Age file

```{r }
#study138data.atri = read_excel("TRCPAD_Participant_Age_20220119.xlsx")

# Download at"
# https://trc.atrihub.org/document_repository/t/transfer-c2n-outbound

#study138data.atri = read.csv("TRCPAD_Participant_Age_20220214.csv")
#study138data.atri = read.csv("TRCPAD_Participant_Age_20220228.csv")
#study138data.atri = read.csv("TRCPAD_Participant_Age_20220502.csv")
study138data.atri = read.csv("TRCPAD_Participant_Age_List.csv")

colnames(study138data.atri) = c("USUBJID","LBREFID", "age")

```

Number of patients in age file

```{r}

length(study138data.atri$LBREFID)

study138data.atri$age = as.numeric(study138data.atri$age)


```



### Differences between age file and MS data

```{r}

if (length(study138data.atri$LBREFID) < length(study138data.c2n$LBREFID)) {
  
  print(paste0("There are less patients in the age file than in the MS file"))
#  knitr::knit_exit()  
  
} else if (length(study138data.atri$LBREFID) > length(study138data.c2n$LBREFID)) {

  print(paste0("There are more patients in the age file than in the MS file"))
#  knitr::knit_exit()  

}
  
```

#### Samples in the C2N set not ATRI set:


LBREFIDs:

```{r}
delta = sort(setdiff(study138data.c2n$LBREFID, study138data.atri$LBREFID))
delta
```

Number of samples:

```{r}
length(setdiff(study138data.c2n$LBREFID, study138data.atri$LBREFID))
```

Number of samples per batch:

```{r}
table(study138data.c2n[which(study138data.c2n$LBREFID %in% delta), c("batch")])
```

#### Samples in ATRI set not in C2N set:

LBREFIDs:

```{r}
sort(setdiff(study138data.atri$LBREFID, study138data.c2n$LBREFID))
```

Number of samples:


```{r}
length(setdiff(study138data.atri$LBREFID, study138data.c2n$LBREFID))

```




## Merge the data

Merge age and mass spec data

```{r }

study138data = merge(study138data.c2n, study138data.atri, all.x = T)

study138data$LBSTAT   = ""
study138data$LBREASND = ""  

```



## Abeta comments:

```{r}
study138data$Abeta_comment[which(study138data$Abeta_comment == "")] = NA
study138data$ApoE_comment[which(study138data$ApoE_comment == "")] = NA
```

```{r}

paste0("Number of comments: ", sum(as.numeric(!is.na(study138data$Abeta_comment))))

formattable(table(study138data$Abeta_comment))

```





## ApoE comments:

```{r}

paste0("Number of comments: ", sum(as.numeric(!is.na(study138data$ApoE_comment))))

formattable(table(study138data$ApoE_comment))

```

## Samples in analysis

For all samples with sample comments - set the results to NA

```{r}

sampleswithqns = which(!is.na(study138data$Abeta_comment) | !is.na(study138data$ApoE_comment))

paste0("Number of samples: ", length(study138data$LBREFID), ", samples with comments: ", length(sampleswithqns))

study138data[sampleswithqns, c("Abeta40_Plasma_IPMS", "Abeta42_Plasma_IPMS", "Abeta4240_Plasma_IPMS", "Proteotype")] = NA

# Set status and reason not done
study138data[sampleswithqns, c("LBSTAT")] = "ND"
study138data[sampleswithqns, c("LBREASND")] = "QNS"

```


# Compute APS on the data

```{r}
# Convert "N/A" to NA
study138data$Abeta40_Plasma_IPMS[which(study138data$Abeta40_Plasma_IPMS == "N/A")] = NA
study138data$Abeta42_Plasma_IPMS[which(study138data$Abeta42_Plasma_IPMS == "N/A")] = NA

# Then convert to numbers
study138data$Abeta40_Plasma_IPMS = as.numeric(study138data$Abeta40_Plasma_IPMS)
study138data$Abeta42_Plasma_IPMS = as.numeric(study138data$Abeta42_Plasma_IPMS)


# Make sure we have complete data for APS calculation
#study138data = study138data[complete.cases(study138data[ , c("Abeta40_Plasma_IPMS", "Abeta40_Plasma_IPMS")]),]


# TODO: setup a function that will do this
study138data$Abeta4240_Plasma_IPMS = study138data$Abeta42_Plasma_IPMS / study138data$Abeta40_Plasma_IPMS

# count apoE4 alleles
study138data$apoe4copynr = 0
study138data$apoe4copynr[is.na(study138data$Proteotype)] = NA
study138data$apoe4copynr[study138data$Proteotype == "E2/E4"] = 1
study138data$apoe4copynr[study138data$Proteotype == "E3/E4"] = 1
study138data$apoe4copynr[study138data$Proteotype == "E4/E4"] = 2
# E4 presence
study138data$apoe4present = (study138data$apoe4copynr > 0)

# count apoE2 alleles
study138data$apoe2copynr = 0
study138data$apoe2copynr[is.na(study138data$Proteotype)] = NA
study138data$apoe2copynr[study138data$Proteotype == "E2/E2"] = 2
study138data$apoe2copynr[study138data$Proteotype == "E2/E3"] = 1
study138data$apoe2copynr[study138data$Proteotype == "E2/E4"] = 1
# E2 presence
study138data$apoe2present = (study138data$apoe2copynr > 0)
```

Samples where we cannot calculate APS:

```{r}

paste0("Samples with no age: ", length(which(is.na(study138data$age))))
paste0("Samples with no apoe4copynr: ", length(which(is.na(study138data$apoe4copynr))))
paste0("Samples with no Abeta4240_Plasma_IPMS: ", length(which(is.na(study138data$Abeta4240_Plasma_IPMS))))

```


```{r}
# Load the model
model = readRDS("model.clia3_25.rds")

# Calculate the response of the model given the new data
study138data$APS = round(predict(model, newdata = study138data, type='response') * 100)

# APS = -1 means no age information or age below 40
study138data$APS[which(is.na(study138data$age))] = -1
study138data$APS[which(study138data$age < 40)] = -1

```

6 random entries for testing APS calculation manually

```{r}

study138data[sample(6), c("LBREFID", "Abeta4240_Plasma_IPMS", "Abeta42_Plasma_IPMS", "Abeta40_Plasma_IPMS", "Proteotype", "age", "APS", "LBREASND")]

```


```{r}
# Assign category
study138data$APS_Category = sapply(study138data$APS, categorize_aps)

table(study138data$APS_Category)

```



# Data review

```{r}

# EXPORT study120data FOR CLIA MODEL
studyid = "STUDY138"
modeldata = study138data[, c("age", "Proteotype", "apoe4copynr", "apoe2copynr", "Abeta42_Plasma_IPMS", "Abeta40_Plasma_IPMS", "Abeta4240_Plasma_IPMS")]

modeldata$race = NA
modeldata$ethnicity = NA
modeldata$gender = NA
modeldata$id = paste0(studyid, "_", study138data$LBREFID)
modeldata$cohort = studyid
#visual read positive - in this case not VR 
modeldata$amyloidpositive = NA
modeldata$centiloid = NA
modeldata$MMSE = NA
modeldata$CDR = NA
modeldata$tracer = "Navidea"
modeldata$diagnosis_desc = NA
modeldata$state = "NA"

write.csv(modeldata, paste0("small_model_data_", studyid, ".csv"))


```



## Demographics

```{r results='asis'}

demographics_table = 
  list(
        "Age" = 
    list("Mean (sd)" = ~ qwraps2::mean_sd(na.omit(.data$age), denote_sd = "paren", digits = 1),
         "Range" = ~ paste0(round(min(na.omit(.data$age)), 1), "-", round(max(na.omit(.data$age)), 1))
         ),

        "ApoE" = 
    list("E2/E2" = ~ paste0(sum(as.numeric(.data$Proteotype == "E2/E2"), na.rm = t), " (", printpct(round(sum(as.numeric(.data$Proteotype == "E2/E2"), na.rm = t)/length(na.omit(.data$Proteotype))*100.0, 1) ), ")"  ) ,
         "E2/E3" = ~ paste0(sum(as.numeric(.data$Proteotype == "E2/E3"), na.rm = t), " (", printpct(round(sum(as.numeric(.data$Proteotype == "E2/E3"), na.rm = t)/length(na.omit(.data$Proteotype))*100.0, 1) ), ")"  ) ,
         "E2/E4" = ~ paste0(sum(as.numeric(.data$Proteotype == "E2/E4"), na.rm = t), " (", printpct(round(sum(as.numeric(.data$Proteotype == "E2/E4"), na.rm = t)/length(na.omit(.data$Proteotype))*100.0, 1) ), ")"  ) ,
         "E3/E3" = ~ paste0(sum(as.numeric(.data$Proteotype == "E3/E3"), na.rm = t), " (", printpct(round(sum(as.numeric(.data$Proteotype == "E3/E3"), na.rm = t)/length(na.omit(.data$Proteotype))*100.0, 1) ), ")"  ) ,
         "E3/E4" = ~ paste0(sum(as.numeric(.data$Proteotype == "E3/E4"), na.rm = t), " (", printpct(round(sum(as.numeric(.data$Proteotype == "E3/E4"), na.rm = t)/length(na.omit(.data$Proteotype))*100.0, 1) ), ")"  ) ,
         "E4/E4" = ~ paste0(sum(as.numeric(.data$Proteotype == "E4/E4"), na.rm = t), " (", printpct(round(sum(as.numeric(.data$Proteotype == "E4/E4"), na.rm = t)/length(na.omit(.data$Proteotype))*100.0, 1) ), ")"  )
         ),

    
          "APS Category" = 
    list("High" = ~ paste0(sum(as.numeric(na.omit(.data$APS_Category) == "High")), " (", printpct(round(sum(as.numeric(na.omit(.data$APS_Category) == "High"))/length(na.omit(na.omit(.data$APS_Category)))*100.0, 1) ), ")"  ) ,
         "Intermediate" = ~ paste0(sum(as.numeric(na.omit(.data$APS_Category) == "Intermediate")), " (", printpct(round(sum(as.numeric(na.omit(.data$APS_Category) == "Intermediate"))/length(na.omit(na.omit(.data$APS_Category)))*100.0, 1) ), ")"  ) ,
         "Low" = ~ paste0(sum(as.numeric(na.omit(.data$APS_Category) == "Low")), " (", printpct(round(sum(as.numeric(na.omit(.data$APS_Category) == "Low"))/length(na.omit(na.omit(.data$APS_Category)))*100.0, 1) ), ")"  ) 
         
         ),


        "Blood Abeta 42/40" = 
    list("mean (sd)" = ~ qwraps2::mean_sd(na.omit(.data$Abeta4240_Plasma_IPMS), denote_sd = "paren", digits = 3),
         "min" = ~ round(min(na.omit(.data$Abeta4240_Plasma_IPMS)), 3),
         "max" = ~ round(max(na.omit(.data$Abeta4240_Plasma_IPMS)), 3)
         ),
    
        "Blood Aβ40" = 
    list("mean (sd)" = ~ qwraps2::mean_sd(na.omit(.data$Abeta40_Plasma_IPMS), denote_sd = "paren", digits = 1),
         "min" = ~ round(min(na.omit(.data$Abeta40_Plasma_IPMS)), digits = 1),
         "max" = ~ round(max(na.omit(.data$Abeta40_Plasma_IPMS)), digits = 1)
         ),
    
        "Blood Aβ42" = 
    list("mean (sd)" = ~ qwraps2::mean_sd(na.omit(.data$Abeta42_Plasma_IPMS), denote_sd = "paren", digits = 1),
         "min" = ~ round(min(na.omit(.data$Abeta42_Plasma_IPMS)), digits = 1),
         "max" = ~ round(max(na.omit(.data$Abeta42_Plasma_IPMS)), digits = 1)
         )
        
  )



# Show data for all subjects
c = summary_table(study138data, demographics_table)

colnames(c) = gsub("x", "-", colnames(c))

c

```


## APS histogram

```{r}
ggplot(study138data, aes(x=APS)) + 
  geom_histogram(aes(y=..density..),fill="dodgerblue3",color="white",alpha=0.7, binwidth = 5) + 
  geom_density() +
  geom_rug() +
  geom_vline(xintercept = 35.5) +
  geom_vline(xintercept = 57.5) +
  labs(x='Abeta4240_Plasma_IPMS') +
  theme_minimal()

```

## Age histogram

```{r}

ggplot(study138data, aes(x=age)) + 
  geom_histogram(aes(y=..density..),fill="dodgerblue3",color="white",alpha=0.7, binwidth = 1) + 
  geom_density() +
  geom_rug() +
  labs(x='Abeta4240_Plasma_IPMS') +
  annotate("text", x=mean(study138data$age), y=0.02, label = paste0("mean = ", round(mean(study138data$age), 3))) +
  theme_minimal()

```



## Abeta histograms

```{r}
ggplot(study138data, aes(x=Abeta4240_Plasma_IPMS)) + 
  geom_histogram(aes(y=..density..),fill="dodgerblue3",color="white",alpha=0.7, binwidth = 0.005) + 
  geom_density() +
  geom_rug() +
  geom_vline(xintercept = 0.089) +
  labs(x='Abeta4240_Plasma_IPMS') +
  annotate("text", x=mean(study138data$Abeta4240_Plasma_IPMS), y=20, label = paste0("mean = ", round(mean(study138data$Abeta4240_Plasma_IPMS), 3))) +
  theme_minimal()

ggplot(study138data, aes(x=Abeta40_Plasma_IPMS)) + 
  geom_histogram(aes(y=..density..),fill="dodgerblue3",color="white",alpha=0.7, binwidth = 50) + 
  geom_density() +
  geom_rug() +
  labs(x='Abeta40_Plasma_IPMS') +
  annotate("text", x=mean(study138data$Abeta40_Plasma_IPMS), y=0.0025, label = paste0("mean = ", round(mean(study138data$Abeta40_Plasma_IPMS), 3))) +
  theme_minimal()

ggplot(study138data, aes(x=Abeta42_Plasma_IPMS)) + 
  geom_histogram(aes(y=..density..),fill="dodgerblue3",color="white",alpha=0.7, binwidth = 5) + 
  geom_density() +
  geom_rug() +
  labs(x='Abeta42_Plasma_IPMS') +
  annotate("text", x=mean(study138data$Abeta42_Plasma_IPMS), y=0.025, label = paste0("mean = ", round(mean(study138data$Abeta42_Plasma_IPMS), 3))) +
  theme_minimal()


```

## Abeta histograms - by batch


```{r}
ggplot(study138data, aes(x=Abeta4240_Plasma_IPMS, color = as.factor(batch))) + 
  geom_histogram(aes(y=..density..),fill="dodgerblue3",color="white",alpha=0.7, binwidth = 0.005) + 
  geom_density() +
  geom_rug() +
  labs(x='Abeta4240_Plasma_IPMS') +
  theme_minimal()

ggplot(study138data, aes(x=Abeta40_Plasma_IPMS, color = as.factor(batch))) + 
  geom_histogram(aes(y=..density..),fill="dodgerblue3",color="white",alpha=0.7, binwidth = 50) + 
  geom_density() +
  geom_rug() +
  labs(x='Abeta40_Plasma_IPMS') +
  theme_minimal()

ggplot(study138data, aes(x=Abeta42_Plasma_IPMS, color = as.factor(batch))) + 
  geom_histogram(aes(y=..density..),fill="dodgerblue3",color="white",alpha=0.7, binwidth = 5) + 
  geom_density() +
  geom_rug() +
  labs(x='Abeta42_Plasma_IPMS') +
  theme_minimal()


```

## Abeta averages tracking by batch


```{r}

ggplot(study138data, aes(x=batch, y=Abeta4240_Plasma_IPMS, color = as.factor(batch))) + 
  geom_boxplot() +
  labs(y='Abeta4240_Plasma_IPMS', x = "Batch") +
  theme_minimal()

ggplot(study138data, aes(x=batch, y=Abeta40_Plasma_IPMS, color = as.factor(batch))) + 
  geom_boxplot() +
  labs(y='Abeta40_Plasma_IPMS', x = "Batch") +
  theme_minimal()

ggplot(study138data, aes(x=batch, y=Abeta42_Plasma_IPMS, color = as.factor(batch))) + 
  geom_boxplot() +
  labs(y='Abeta42_Plasma_IPMS', x = "Batch") +
  theme_minimal()

```



# Get data ready for export



## Convert data from wide to long format

Test codes:

+---------------+------------------------------------+----------------+
| LBTESTCD      | LBTEST                             | LBORRESU       |
+===============+====================================+================+
| AB42          | PLASMA ABETA42                     | pg/mL          |
+---------------+------------------------------------+----------------+
| AB40          | PLASMA ABETA40                     | pg/mL          |
+---------------+------------------------------------+----------------+
| AB4240        | ABETA42/40 RATIO                   |                |
|               |                                    |                |
|               | Provided as 3 decimal point        |                |
|               | precision (eg. 0.090)              |                |
+---------------+------------------------------------+----------------+
| APS           | AMYLOID PROBABILITY SCORE          |                |
|               |                                    |                |
|               | Integer between 0 and 100 (both    |                |
|               | inclusive)                         |                |
+---------------+------------------------------------+----------------+
| APOE          | APOE PROTEOTYPE                    |                |
+---------------+------------------------------------+----------------+

```{r}

# 1) Keep just the data we report
study138data.export = study138data[, c("USUBJID", "LBREFID", "LBSTAT", "LBREASND", "Abeta40_Plasma_IPMS", "Abeta42_Plasma_IPMS", "Abeta4240_Plasma_IPMS", "APS", "Proteotype")]
colnames(study138data.export) = c("USUBJID", "LBREFID", "LBSTAT", "LBREASND", "AB40", "AB42", "AB4240", "APS", "APOE")

# 1.1) Round the ratio to 3 sig digits
study138data.export$AB4240 = round(study138data.export$AB4240, 3)

# 1.2) GENOPYE SHOULD BE _ not /
study138data.export$APOE = gsub('/', '_', study138data.export$APOE)

# 2) Convert to long format
study138data.export.long = melt(study138data.export, id.vars = c("USUBJID","LBREFID", "LBSTAT", "LBREASND"), variable.name = "LBTESTCD", value.name = "LBORRES")

# 3) Sort by USUBJID & LBREFID
study138data.export.long = study138data.export.long[
  with(study138data.export.long, order(USUBJID, LBREFID)),
]

formattable(head(study138data.export.long, n = 10))

# For samples missing age data (APS = -1) - fill in LBSTAT, LBREASND then change LBORRES to ""
study138data.export.long[which(study138data.export.long$LBORRES == -1 & study138data.export.long$LBSTAT == "" & study138data.export.long$LBTESTCD == "APS"), "LBSTAT"] = "ND"
study138data.export.long[which(study138data.export.long$LBORRES == -1 & study138data.export.long$LBREASND == "" & study138data.export.long$LBTESTCD == "APS"), "LBREASND"] = "Patient age missing or low"
study138data.export.long[which(study138data.export.long$LBORRES == -1 & study138data.export.long$LBTESTCD == "APS"), "LBORRES"] = ""


# For QNS samples all the results are NA - but they do not want NA in the LBORRES column so we will change all NA to ""
study138data.export.long[which(is.na(study138data.export.long$LBORRES)), "LBORRES"] = "" 
  
```

## Add the other fields to the export file

Order of fields in the output file:

| Field Name | Value       |
|------------|-------------|
| STUDYID    | "TRCPAD"    |
| DOMAIN     | "LB"        |
| SITE       | ""          |
| USUBJID    | dynamic     |
| LBNAM      | "C2N"       |
| LBREFID    | dynamic     |
| VISIT      | "sc1"       |
| LBDTC      | ""          |
| LBSPEC     | "PLASMA"    |
| LBCAT      | "BIOMARKER" |
| LBTESTCD   | dynamic     |
| LBORRES    | dynamic     |
| LBORRESU   | per above   |
| LBMETHOD   | "LC-MS/MS"  |
| LBSTAT     | dynamic     |
| LBREASND   | dynamic     |


```{r}
# STATIC ONES FIRST:
study138data.export.long$STUDYID  = "TRCPAD"
study138data.export.long$DOMAIN   = "LB"
study138data.export.long$SITE     = ""
study138data.export.long$LBNAM    = "C2N" 
study138data.export.long$VISIT    = "sc1"
study138data.export.long$LBDTC    = ""
study138data.export.long$LBSPEC   = "PLASMA"  
study138data.export.long$LBCAT    = "BIOMARKER"
study138data.export.long$LBMETHOD = "LC-MS/MS"
#study138data.export.long$LBSTAT   = ""
#study138data.export.long$LBREASND = ""  

# UNIT DEPENDS ON LBTESTCD
study138data.export.long$LBORRESU = "" 
study138data.export.long$LBORRESU[which(study138data.export.long$LBTESTCD == "AB42" | study138data.export.long$LBTESTCD == "AB40")] = "pg/mL"



```


```{r}

# ORDER THE COLUMNS
col_order <- c("STUDYID", "DOMAIN", "SITE", "USUBJID", "LBNAM", "LBREFID", "VISIT", "LBDTC", "LBSPEC", "LBCAT", "LBTESTCD", "LBORRES", "LBORRESU", "LBMETHOD", "LBSTAT", "LBREASND")
study138data.export.long.sort = study138data.export.long[, ..col_order]

formattable(head(study138data.export.long.sort, n = 10), row.names=FALSE)
  
```



Number of samples in final file:

```{r}

length(study138data$LBREFID)

```

Number of lines in the long export file divided by 5 since there are 5 results per sample.

```{r}

length(study138data.export.long.sort$USUBJID)/5

```



```{r}

# Filenaming convention:

## TEST
# TRCPAD_TEST_C2N_BM_FULL_YYYYMMDDX.csv

## PROD
# TRCPAD_PROD_C2N_BM_FULL_YYYYMMDDX.csv

currentDate = format(Sys.Date(), "%Y%m%d") 
filename = paste0("TRCPAD_PROD_C2N_BM_FULL_", currentDate, "1.csv")

write.csv(study138data.export.long.sort, filename, row.names=FALSE)

```


```{r message=FALSE}

message(paste0("Data export saved to: ", filename))
message(paste0("Time to run this code: ", difftime(Sys.time(), now)))

```


