---
title: "STUDY160 - Interim Unaudited Analysis"
author: "C2N Diagnostics - proprietary and confidential information. Not to be copied or distributed without C2N written consent."
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)

now <<- Sys.time()

library(rmarkdown)
library(ggplot2)
library(pROC)
library(data.table)
library(caret)
library(plyr)
library(dplyr)
library(ggpubr)
library(expss)
library(ggforce)
library(gridExtra)
# visualizing the effects of variables in a GLM
library(effects)
library(Matrix)

# For nicer looking html tables
library(formattable)
library(readxl)

# For summary tables
library(qwraps2)
options(qwraps2_markup = "markdown")

library(DiagrammeR)
#http://rich-iannone.github.io/DiagrammeR/graphviz_and_mermaid.html
#https://bookdown.org/yihui/rmarkdown-cookbook/diagrams.html




# Plot & visualization
theme_new <- function(base_size = 12, base_family = ""){
  theme_bw(base_size = base_size, base_family = base_family) %+replace%
    theme(
      legend.key=element_rect(colour=NA, fill =NA),
      panel.grid = element_blank(),   
      panel.border = element_rect(fill = NA, colour = "black", size=1),
      panel.background = element_rect(fill = "white", colour = "black"), 
      strip.background = element_rect(fill = NA)
    )
}

theme_set(theme_new(base_size = 16))


printpct = function(percentage) {
  percentage = as.numeric(percentage)
  
  if (is.na(percentage) | percentage == 0)
    return("-")
  else
    return(paste0(percentage, "%"))
}

categorize_aps = function (APS) {
  
  if (is.na(APS)) {
    return(NA)
  }
  
  if (APS <= 35) {
    return("Low")
  } else if (APS >= 58) {
    return("High")
  } else {
    return("Intermediate")
  }
  
}


```


```{r}
abetadata = read_xlsx("../Lab Data/Study160_Combined Abeta Data.xlsx", col_types = c("text"))
colnames(abetadata) = c("Batch", "Accession", "Abeta_40", "Abeta_42", "ratio_excel", "Comments")

# convert results to numbers
abetadata$Abeta_40 = as.numeric(abetadata$Abeta_40)
abetadata$Abeta_42 = as.numeric(abetadata$Abeta_42)
abetadata$Abeta_ratio = abetadata$Abeta_42 / abetadata$Abeta_40


manifest = read_xlsx("../Sample Management/manifest for sarah-clean.xlsx")
colnames(manifest) = c("Study", "SampleID", "Patient_ID", "Accession", "Patient_number", "Position", "Notes")

study160data = merge(abetadata, manifest, all.x = T)

study160data = study160data[, c("Study", "Patient_ID", "SampleID", "Accession", "Abeta_40", "Abeta_42", "Abeta_ratio")]

paged_table(study160data)
```

They mixed accession numbers and sample numbers in the manifest and in the Abeta file - we need to fix this

```{r}

paste0("Number of data points: ", length(abetadata$Accession))

# 1) Merge in based on accession number
study160data.accessionnumber = merge(abetadata, manifest)
study160data.accessionnumber = study160data.accessionnumber[, c("Study", "Patient_ID", "SampleID", "Accession", "Abeta_40", "Abeta_42", "Abeta_ratio")]


# 2) Merge in based on SampleID
manifest2 = manifest
colnames(manifest2) = c("Study", "Accession", "Patient_ID", "SampleID", "Patient_number", "Position", "Notes")
study160data.sampleid = merge(abetadata, manifest2)
study160data.sampleid = study160data.sampleid[, c("Study", "Patient_ID", "SampleID", "Accession", "Abeta_40", "Abeta_42", "Abeta_ratio")]

paste0("Combined datapoints from two independent merges: ", length(study160data.accessionnumber$Accession) + length(study160data.sampleid$Accession))

# 3) combine and remove duplicates that
study160data = rbind(study160data.accessionnumber, study160data.sampleid)
study160data = study160data[!duplicated(study160data), ]

paste0("Number of unique points after merge with manifest and removing duplicates: ", length(study160data$Accession))

if (length(study160data$Accession) != length(abetadata$Accession)) {

  print("Merge did not work")
    
  knitr::knit_exit()
  
}




```

How did merge go? List of samples with missing patient id:

```{r}

study160data[which(is.na(study160data$Patient_ID)), ]

```

Prepare data for export


Study
Patient ID
Abeta 40 
Abeta 40 units
Abeta 42
Abeta 42 units 
Abeta ratio 
Phosphorylated Tau181
Phosphorylated Tau181 units
non-phosphorylated Tau181  
non-phosphorylated Tau181 units
p-Tau181/non-p-Tau181 ratio
Phosphorylated Tau217 
Phosphorylated Tau217 units 
non-phosphorylated Tau217 
non-phosphorylated Tau217 units 
p-Tau217/non-p- Tau181 ratio

```{r}

study160data$Abeta_40_units = "pg/mL"
study160data$Abeta_42_units = "pg/mL"

study160data$ptau181 = "" 
study160data$ptau181_units = "pg/mL" 
study160data$nptau181 = "" 
study160data$nptau181_units = "pg/mL" 
study160data$ptau181ratio = "" 

study160data$ptau217 = "" 
study160data$ptau217_units = "pg/mL" 
study160data$nptau217 = "" 
study160data$nptau217_units = "pg/mL" 
study160data$ptau217ratio = "" 



study160data.export = study160data[, c("Study", "Patient_ID", "Abeta_40", "Abeta_40_units", "Abeta_42", "Abeta_40_units", "Abeta_ratio", "ptau181", "ptau181_units", "nptau181", "nptau181_units", "ptau181ratio", "ptau217", "ptau217_units", "nptau217", "nptau217_units", "ptau217ratio")]

colnames(study160data.export) = c("Study", "Patient ID", "Abeta 40 ", "Abeta 40 units", "Abeta 42", "Abeta 42 units ", "Abeta ratio ", "Phosphorylated Tau181", "Phosphorylated Tau181 units", "non-phosphorylated Tau181  ", "non-phosphorylated Tau181 units", "p-Tau181/non-p-Tau181 ratio", "Phosphorylated Tau217 ", "Phosphorylated Tau217 units ", "non-phosphorylated Tau217 ", "non-phosphorylated Tau217 units ", "p-Tau217/non-p- Tau181 ratio")

write.csv(study160data.export, "C2N_Biomarkers.csv", row.names=FALSE)

```




